---
title: "STATS 604 - Project 4 EDA"
output: html_document
date: "`r Sys.Date()`"
---

Analyze midpoint temps to see if it’s a reasonable replacement for average temp.
Make scatterplots by prediction station, time of year to make sure relationship is always reasonably strong.

```{r setup, include=FALSE}
library(knitr)
library(lubridate)
library(tidyverse)
opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE}
data <- dir("data/raw/stations/", full.names = TRUE) %>%
  map_dfr(read_csv, id = "station2", show_col_types = FALSE) %>%
  mutate(station2 = map_chr(station2, ~ str_sub(basename(.x), 1, 4)))
```

Some stations have far fewer rows than the others.
```{r}
data %>%
  count(station2, sort = TRUE) %>%
  kable(
    col.names = c("Station", "Number of Rows"),
    format.args = list(big.mark = ",")
  )
```

KDEN and KMIC are missing observations for most years.
```{r}
data %>%
  mutate(
    date_year = year(date), date_month = month(date), date_day = day(date)
  ) %>%
  group_by(station2, date_month, date_day) %>%
  summarize(num_years = n_distinct(date_year)) %>%
  ungroup() %>%
  mutate(date = make_date(month = date_month, day = date_day)) %>%
  ggplot(aes(date, num_years)) +
  facet_wrap(vars(station2)) +
  geom_line() +
  labs(x = "Date", y = "Number Of Years Date Is In") +
  theme_bw()
```

KDEN and KMIC data don't go back as far as the data for the other stations.
```{r}
data %>%
  filter(station2 %in% c("KDEN", "KMIC")) %>%
  group_by(station2) %>%
  summarize(min_date = min(date))
```


```{r}
data %>%
  mutate(date_year = update(date, months = 1, days = 1)) %>%
  select(where(is.double) & !date) %>%
  group_by(date_year) %>%
  summarize(across(.fns = ~ mean(is.na(.x)))) %>%
  pivot_longer(cols = !date_year, names_to = "var", values_to = "prop_na") %>%
  ggplot(aes(date_year, prop_na)) +
  facet_wrap(vars(var)) +
  geom_line() +
  labs(x = "Year", y = "Proportion Missing") +
  theme_bw()
```

```{r}
data %>%
  mutate(date_year = update(date, months = 1, days = 1)) %>%
  select(station2, where(is.double) & !date) %>%
  group_by(station2, date_year) %>%
  summarize(across(.fns = ~ mean(is.na(.x)))) %>%
  ungroup() %>%
  pivot_longer(
    cols = !c(station2, date_year), names_to = "var", values_to = "prop_na"
  ) %>%
  ggplot(aes(date_year, prop_na, color = station2)) +
  facet_wrap(vars(var)) +
  geom_line() +
  labs(x = "Year", y = "Proportion Missing") +
  theme_bw()
```

```{r}
data %>%
  mutate(midpoint = (TMIN + TMAX) / 2) %>%
  ggplot(aes(midpoint, TAVG)) +
  facet_wrap(vars(station2)) +
  geom_point() +
  theme_bw()
```

